---
title: "Knot Spacing and Bandwidth Sim"
author: "Morris, S."
date: "July 6, 2015"
output: pdf_document
---

```{r initial-setup, cache = TRUE, echo = FALSE, include = FALSE}
library(fields)
library(evd)
library(spBayes)
library(fields)
library(SpatialTools)
# library(microbenchmark)
library(mvtnorm)
library(Rcpp)
library(numDeriv)
Sys.setenv("PKG_CXXFLAGS"="-fopenmp")
Sys.setenv("PKG_LIBS"="-fopenmp")
sourceCpp(file = "../../code/R/pairwise.cpp")

source("../../code/R/auxfunctions.R", chdir = TRUE)
source("../../code/R/updateModel.R")
source("../../code/R/mcmc.R")
source("../../code/R/probit.R", chdir=T)
```

# Small-scale simulation study on knots and bandwidth

We tried two different true knot spacings in this simulation study.
The first spacing sets knots on a $20 \times 20$ grid in $[0, 1] \times [0, 1]$.
The second spacing sets knots on a $10 \times 10$ grid in $[0, 1] \times [0, 1]$.
We used three different data generation settings.

1. $\rho = 0.01$ and $\pi = 0.01$ with $n_s = 2000$ sites.
2. $\rho = 0.01$ and $\pi = 0.05$ with $n_s = 1000$ sites.
3. $\rho = 0.02$ and $\pi = 0.05$ with $n_s = 1000$ sites.

These all set $\rho$ to be the knot spacing used in fitting the GEV which is a $12 \times 12$ grid.

## Setting 1: $\rho = 0.01$ and $\pi = 0.01$ with $n_s = 2000$ sites.

```{r gendata-1-1, cache = TRUE, echo = FALSE, include = FALSE}
#### when rho = 0.01 and prop = 0.01
set.seed(7483)  # site
ns    <- 2000
s     <- cbind(runif(ns), runif(ns))
knots.t1 <- expand.grid(seq(0, 1, length=20), seq(0, 1, length=20))
knots.t2 <- expand.grid(seq(0, 1, length=10), seq(0, 1, length=10))
x     <- matrix(1, ns, 1)

alpha.t <- 0.3
rho.t   <- 0.01
xi.t    <- 0
prop.t  <- 0.01

set.seed(3282)  # data
nreps <- 10
nsettings <- 2
y <- matrix(data = NA, nrow = ns, ncol = nreps * nsettings)
thresh <- rep(NA, nreps * nsettings)
for (i in 1:nreps) {
  idx <- (i - 1) * nsettings + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t1, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh

  idx <- idx + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t2, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh
}

# note, the way the data are generated, it means every other data set has the same knot spacing
```

```{r results-1-1-20, cache = TRUE, echo = FALSE, include = FALSE}
bs1.1.20 <- matrix(NA, 3, 10)
for (i in 1:10) {
  set <- (i - 1) * 2 + 1
  file <- paste("sim-results/sim-knots-", set, "-1-1.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs1.1.20[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs1.1.20[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs1.1.20[3, i] <- BrierScore(post.prob.pro, y.i.p)
}
```

```{r results-1-1-10, cache = TRUE, echo = FALSE, include = FALSE}
bs1.1.10 <- matrix(NA, 3, 10)
idx <- 1
for (i in 1:10) {
  set <- i * 2
  file <- paste("sim-results/sim-knots-", set, "-1-1.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs1.1.10[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs1.1.10[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs1.1.10[3, i] <- BrierScore(post.prob.pro, y.i.p)
  idx <- idx + 1
}
```

\scriptsize
```{r simresults-1-1, echo = FALSE}
library(pander)
panderOptions("digits", 4)
rownames(bs1.1.20) <- rownames(bs1.1.10) <- c("gev", "log", "pro")
bs1.1.20 <- cbind(bs1.1.20, rowMeans(bs1.1.20, na.rm = TRUE))
bs1.1.10 <- cbind(bs1.1.10, rowMeans(bs1.1.10, na.rm = TRUE))
colnames(bs1.1.20) <- colnames(bs1.1.10) <- c(1:10, "Mean")
set.caption("Results (x 100) for Knots: 20 x 20, $\\rho = 0.01$, $\\pi = 0.01$")
pander(t(bs1.1.20) * 100)

set.caption("Results (x 100) for Knots: 10 x 10, $\\rho = 0.01$, $\\pi = 0.01$")
pander(t(bs1.1.10) * 100)
```

## Setting 2: $\rho = 0.01$ and $\pi = 0.05$ with $n_s = 1000$ sites.

```{r gendata-5-1, cache = TRUE, echo = FALSE, include = FALSE}
#### when rho = 0.01 and prop = 0.05
set.seed(7483)  # site
ns    <- 1000
s     <- cbind(runif(ns), runif(ns))
knots.t1 <- expand.grid(seq(0, 1, length=20), seq(0, 1, length=20))
knots.t2 <- expand.grid(seq(0, 1, length=10), seq(0, 1, length=10))
x     <- matrix(1, ns, 1)

alpha.t <- 0.3
rho.t   <- 0.01
xi.t    <- 0
prop.t  <- 0.05

set.seed(3282)  # data
nreps <- 10
nsettings <- 2
y <- matrix(data = NA, nrow = ns, ncol = nreps * nsettings)
thresh <- rep(NA, nreps * nsettings)
for (i in 1:nreps) {
  idx <- (i - 1) * nsettings + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t1, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh

  idx <- idx + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t2, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh

  idx <- idx + 1
}
```

```{r results-5-1-20, cache = TRUE, echo = FALSE, include = FALSE}
bs5.1.20 <- matrix(NA, 3, 10)
for (i in 1:10) {
  set <- (i - 1) * 2 + 1
  file <- paste("sim-results/sim-knots-", set, "-5-1.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs5.1.20[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs5.1.20[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs5.1.20[3, i] <- BrierScore(post.prob.pro, y.i.p)
}
```

```{r results-5-1-10, cache = TRUE, echo = FALSE, include = FALSE}
bs5.1.10 <- matrix(NA, 3, 10)
idx <- 1
for (i in 1:10) {
  set <- i * 2
  file <- paste("sim-results/sim-knots-", set, "-5-1.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs5.1.10[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs5.1.10[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs5.1.10[3, i] <- BrierScore(post.prob.pro, y.i.p)
  idx <- idx + 1
}
```

\scriptsize
```{r simresults-5-1, echo = FALSE}
library(pander)
panderOptions("digits", 4)
rownames(bs5.1.20) <- rownames(bs5.1.10) <- c("gev", "log", "pro")
bs5.1.20 <- cbind(bs5.1.20, rowMeans(bs5.1.20, na.rm = TRUE))
bs5.1.10 <- cbind(bs5.1.10, rowMeans(bs5.1.10, na.rm = TRUE))
colnames(bs5.1.20) <- colnames(bs5.1.10) <- c(1:10, "Mean")
set.caption("Results (x 100) for Knots: 20 x 20, $\\rho = 0.01$, $\\pi = 0.05$")
pander(t(bs5.1.20) * 100)

set.caption("Results (x 100) for Knots: 10 x 10, $\\rho = 0.01$, $\\pi = 0.05$")
pander(t(bs5.1.10) * 100)
```

## Setting 3: $\rho = 0.02$ and $\pi = 0.05$ with $n_s = 1000$ sites.

```{r gendata-5-2, cache = TRUE, echo = FALSE, include = FALSE}
#### when rho = 0.02 and prop = 0.05
set.seed(7483)  # site
ns    <- 1000
s     <- cbind(runif(ns), runif(ns))
knots.t1 <- expand.grid(seq(0, 1, length=20), seq(0, 1, length=20))
knots.t2 <- expand.grid(seq(0, 1, length=10), seq(0, 1, length=10))
x     <- matrix(1, ns, 1)

alpha.t <- 0.3
rho.t   <- 0.02
xi.t    <- 0
prop.t  <- 0.05

set.seed(3282)  # data
nreps <- 10
nsettings <- 2
y <- matrix(data = NA, nrow = ns, ncol = nreps * nsettings)
thresh <- rep(NA, nreps * nsettings)
for (i in 1:nreps) {
  idx <- (i - 1) * nsettings + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t1, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh

  idx <- idx + 1
  data <- rRareBinarySpat(x, s = s, knots = knots.t2, beta = 0, xi = xi.t,
                          alpha = alpha.t, rho = rho.t, prob.success = prop.t)

  y[, idx] <- data$y
  thresh[idx] <- data$thresh

  idx <- idx + 1
}
```

```{r results-5-2-20, cache = TRUE, echo = FALSE, include = FALSE}
bs5.2.20 <- matrix(NA, 3, 10)
for (i in 1:10) {
  set <- (i - 1) * 2 + 1
  file <- paste("sim-results/sim-knots-", set, "-5-2.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs5.2.20[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs5.2.20[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs5.2.20[3, i] <- BrierScore(post.prob.pro, y.i.p)
}
```

```{r results-5-2-10, cache = TRUE, echo = FALSE, include = FALSE}
bs5.2.10 <- matrix(NA, 3, 10)
idx <- 1
for (i in 1:10) {
  set <- i * 2
  file <- paste("sim-results/sim-knots-", set, "-5-2.RData", sep = "")
  print(paste("start: set", set))
  load(file)
  bs5.2.10[1, i] <- BrierScore(post.prob.gev, y.i.p)
  bs5.2.10[2, i] <- BrierScore(post.prob.log, y.i.p)
  bs5.2.10[3, i] <- BrierScore(post.prob.pro, y.i.p)
  idx <- idx + 1
}
```

\scriptsize
```{r simresults-5-2, echo = FALSE}
library(pander)
panderOptions("digits", 4)
rownames(bs5.2.20) <- rownames(bs5.2.10) <- c("gev", "log", "pro")
bs5.2.20 <- cbind(bs5.2.20, rowMeans(bs5.2.20, na.rm = TRUE))
bs5.2.10 <- cbind(bs5.2.10, rowMeans(bs5.2.10, na.rm = TRUE))
colnames(bs5.2.20) <- colnames(bs5.2.10) <- c(1:10, "Mean")
set.caption("Results (x 100) for Knots: 20 x 20, $\\rho = 0.02$, $\\pi = 0.05$")
pander(t(bs5.2.20) * 100)

set.caption("Results (x 100) for Knots: 10 x 10, $\\rho = 0.02$, $\\pi = 0.05$")
pander(t(bs5.2.10) * 100)
```

## Next steps

Fit a few datasets with $rho$ as a parameter in the MCMC.