---
title: "Pairwise Composite Likelihood"
author: "Sam Morris"
date: "18 May 2015"
output: pdf_document
---

```{r setup, echo=FALSE, include=FALSE}
# libraries
library(fields)
library(Rcpp)
library(evd)
Sys.setenv("PKG_CXXFLAGS"="-fopenmp")
Sys.setenv("PKG_LIBS"="-fopenmp")
sourceCpp(file = "../R/pairwise.cpp")

source("../R/auxfunctions.R", chdir = TRUE)
source("../R/updateModel.R")
source("../R/mcmc.R")

# knots and locations
set.seed(7483)  # site
ns    <- 200
s     <- cbind(runif(ns), runif(ns))
knots <- expand.grid(seq(0.01, 0.99, length=12), seq(0.01, 0.99, length=12))
dw2   <- rdist(s, knots)
x     <- matrix(1, ns, 1)

# grid for searching over alpha and rho
alphas <- seq(0.1, 0.9, by=0.05)
rhos   <- (knots[2, 1] - knots[1, 1]) * seq(1, 3, by = 0.5)

# plotting variables
xplot <- rep(alphas, each=length(rhos))
yplot <- rep(rhos, length(alphas))
color <- two.colors(n = 256, start = "dodgerblue4", end = "firebrick4", 
                    middle = "white")
```

## Likelihood for the data

The bivariate likelihood for the data is given as

\begin{align}
  f(Y_1, Y_2) = \left\{ \begin{array}{ll}
    1 - \exp \left\{ - \frac{ 1 }{ z_1 } \right\} - \exp \left\{ - \frac{ 1 }{ z_2 } \right\} + \exp \left\{ - \vartheta(\mbox{\bf s}_1, \mbox{\bf s}_2) \right\} \quad & Y_1 = 1, Y_2 = 1 \\[0.5em]
    \exp \left\{ - \frac{ 1 }{ z_2 } \right\} - \exp \left\{ - \vartheta(\mbox{\bf s}_1, \mbox{\bf s}_2) \right\} \quad & Y_1 = 1, Y_2 = 0 \\[0.5em]
    \exp \left\{ - \frac{ 1 }{ z_1 } \right\} - \exp \left\{ - \vartheta(\mbox{\bf s}_1, \mbox{\bf s}_2) \right\} \quad & Y_1 = 0, Y_2 = 1 \\[0.5em]
    \exp \left\{ -\vartheta(\mbox{\bf s}_1, \mbox{\bf s}_2) \right\} \quad & Y_1 = 0, Y_2 = 0
  \end{array} \right.
\end{align}

where $z_i = \left(1 - \xi \mbox{\bf X}_i^T \beta \right)^{1 / \xi}$, and $\vartheta(\mbox{\bf s}_1, \mbox{\bf s}_2) = \sum_{ l = 1 }^{ L } \left[ \left( \frac{ w_l (\mbox{\bf s}_1 ) }{ z_1 } \right)^{1/\alpha} + \left( \frac{ w_{l }(\mbox{\bf s}_2) }{ z_2 } \right)^{1/\alpha} \right]^\alpha$
 

## Pairwise composite likelihood

We initially tried using `optim` on the likelihood with four parameters: a) $\alpha$, b) $\rho$, c) $\xi$, and d) $\beta$, but it appears to be challenging for `optim` to find the true values.
So, we decided to make a grid of $\alpha$ and $\rho$ and run `optim` to only fit $\xi$ and $\beta$

## Data settings

Right now, we're fitting $n = 200$ observations with one replication. In the future, we plan to allow for multiple replications.
In each plot, we give the optimized value for the negative log likelihood when conducting a grid search.
To examine accuracy, the true values for $\alpha$ and $\rho$ are given by the circle, and the optimized values for $\alpha$ and $\rho$ are given by the triangle.

```{r setting1, echo=FALSE, include=FALSE}
# settings for results 1 and results 2
alpha.t <- 0.3
rho.t   <- 0.1
xi.t    <- 0.25
prop    <- c(0.05, 0.01)
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$
```{r fit1, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results1 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results1[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```


```{r plot1, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results1, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results1 == min(results1), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit2, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results2 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results2[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot2, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results2, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results2 == min(results2), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```


```{r setting2, echo=FALSE, include=FALSE}
# settings for results 3 and results 4
alpha.t <- 0.3
rho.t   <- 0.2
xi.t    <- 0.25
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$

```{r fit3, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results3 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results3[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot3, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results3, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results3 == min(results3), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit4, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results4 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results4[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot4, echo=FALSE}
par(mfrow = c(1, 2))
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results4, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results4 == min(results4), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

```{r setting3, echo=FALSE, include=FALSE}
# settings for results 5 and results 6
alpha.t <- 0.5
rho.t   <- 0.1
xi.t    <- 0.25
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$

```{r fit5, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results5 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results5[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot5, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results5, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results5 == min(results5), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit6, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results6 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results6[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot6, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results6, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results6 == min(results6), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

```{r setting4, echo=FALSE, include=FALSE}
# settings for results 7 and results 8
alpha.t <- 0.5
rho.t   <- 0.2
xi.t    <- 0.25
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$

```{r fit7, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results7 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results7[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot7, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results7, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results7 == min(results7), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit8, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results8 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results8[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot8, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results8, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results8 == min(results8), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

```{r setting5, echo=FALSE, include=FALSE}
# settings for results 9 and results 10
alpha.t <- 0.7
rho.t   <- 0.1
xi.t    <- 0.25
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$

```{r fit9, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results9 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results9[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot9, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results9, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results9 == min(results9), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit10, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results10 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results10[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot10, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results10, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results10 == min(results10), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

```{r setting6, echo=FALSE, include=FALSE}
# settings for results 11 and results 12
alpha.t <- 0.7
rho.t   <- 0.2
xi.t    <- 0.25
```

### $\alpha = `r alpha.t`, \pi = `r prop[1]`, \rho = `r rho.t`$

```{r fit11, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[1])
y <- data$y
results11 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results11[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot11, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results11, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[1]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results11 == min(results11), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```

### $\alpha = `r alpha.t`, \pi = `r prop[2]`, \rho = `r rho.t`$

```{r fit12, echo=FALSE, include=FALSE, cache=TRUE}
set.seed(3282)  # data
data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t,
                        alpha = alpha.t, rho = rho.t, prob.success = prop[2])
y <- data$y
results12 <- matrix(9999999, length(rhos), length(alphas))
for (j in seq_along(alphas)) {
  for (i in seq_along(rhos)) {
    results12[i, j] <- fit.rarebinaryCPP(c(0, -4), alpha = alphas[j], 
                                        rho = rhos[i],  y = y, dw2 = dw2,
                                        cov = x, threads = 1)$value
    print(paste("alpha = ", alphas[j], ", rho = ", round(rhos[i], 3), sep=""))
  }
}
```

```{r plot12, echo=FALSE}
par(mfrow = c(1, 2))
plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
     main = "simulated dataset", xlab="", ylab="")
points(s[which(y != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
points(s[which(y == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")

quilt.plot(x = xplot, y = yplot, z = results12, 
           nx = length(alphas), ny = length(rhos), col = color,
           main=bquote(paste("nll grid search (", pi==.(prop[2]), ")")), 
           xlab=bquote(alpha), ylab=bquote(rho))
points(alpha.t, rho.t, col = "gray16", bg = "gray60",
       pch = 21, cex = 1.4)
min.idx <- which(results12 == min(results12), arr.ind = TRUE)
points(alphas[min.idx[2]], rhos[min.idx[1]], col = "gray16", bg = "gray60",
       pch = 24, cex = 1.4)
```