---
title: "Impact of shape parameter"
author: "Sam Morris"
date: "23 June 2015"
output: pdf_document
---

```{r setup, echo=FALSE, include=FALSE}
# libraries
library(fields)
library(Rcpp)
library(evd)
Sys.setenv("PKG_CXXFLAGS"="-fopenmp")
Sys.setenv("PKG_LIBS"="-fopenmp")
sourceCpp(file = "../code/R/pairwise.cpp")

source("../code/R/auxfunctions.R", chdir = TRUE)
source("../code/R/updateModel.R")
source("../code/R/mcmc.R")

# knots and locations
set.seed(7483)  # site
ns    <- 4000
s     <- cbind(runif(ns), runif(ns))
knots <- expand.grid(seq(0.01, 0.99, length=12), seq(0.01, 0.99, length=12))
dw2   <- rdist(s, knots)
x     <- matrix(1, ns, 1)
```

```{r setting1, echo=FALSE, include=FALSE}
# settings for results 1 and results 2
alpha.t <- c(0.2, 0.8)
rho.t   <- 0.1
xi.t    <- c(-0.5, 0.0, 0.5)
prop    <- c(0.01, 0.05)
```

# Data settings:

We are exploring how $\xi$ impacts the realizations.
For each of the following plots, the open circles are the knot locations, the red circles represent a location where the response is 1, and the blue circles represent a location where the response is 0.

\newpage

## Strong spatial dependence: $\alpha = `r alpha.t[1]`$ with $\pi = `r prop[1]`$

### Sample datasets

```{r data1-1, echo=FALSE, include=FALSE, cache=FALSE}
set.seed(3282)  # data
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```

```{r plot1-1, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data1-2, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```


```{r plot1-2, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data1-3, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```


```{r plot1-3, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

## Strong spatial dependence: $\alpha = `r alpha.t[1]`$ with $\pi = `r prop[2]`$

### Sample datasets

```{r data2-1, echo=FALSE, include=FALSE, cache=FALSE}
set.seed(3282)  # data
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```

```{r plot2-1, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data2-2, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```


```{r plot2-2, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data2-3, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[1], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```


```{r plot2-3, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

## Weak spatial dependence: $\alpha = `r alpha.t[2]`$ with $\pi = `r prop[1]`$

### Sample datasets

```{r data3-1, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```

```{r plot3-1, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data3-2, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```

```{r plot3-2, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data3-3, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[1])
  y[, i] <- data$y
}
```

```{r plot3-3, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

## Weak spatial dependence: $\alpha = `r alpha.t[2]`$ with $\pi = `r prop[2]`$

### Sample datasets

```{r data4-1, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```

```{r plot4-1, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data4-2, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```

```{r plot4-2, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```

```{r data4-3, echo=FALSE, include=FALSE, cache=FALSE}
y <- matrix(NA, ns, 3)
for (i in 1:3) {
  data <- rRareBinarySpat(x, s = s, knots = knots, beta = 0, xi = xi.t[i],
                        alpha = alpha.t[2], rho = rho.t, prob.success = prop[2])
  y[, i] <- data$y
}
```

```{r plot4-3, echo=FALSE, cache=FALSE, fig.height=3, fig.width=7}
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(knots, ylim = c(0, 1), xlim = c(0, 1), 
       main = bquote(paste("simulated dataset: ", xi ==.(xi.t[i]))), xlab = "", ylab = "")
  points(s[which(y[, i] != 1), ], pch = 21, col = "dodgerblue4", bg = "dodgerblue1")
  points(s[which(y[, i] == 1), ], pch = 21, col = "firebrick4", bg = "firebrick1")
}
```